---
title: "Homework Markdown Document"
author: "Danielle Barnas and Group 1"
date: "3/3/2021"
output: 
  prettydoc::html_pretty:
    toc: true
    theme: hpstr
---

```{r setup, include=FALSE, fig.path = "../Output/"}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Introduction
This document will display figures and information from submarine groundwater data taken in Hawaii by Dr. Nyssa Silbiger and her team.

# Load Libraries
```{r}
library(tidyverse)
library(here)
library(kableExtra)
library(PNWColors)
```

# Data Analysis

### Bring in data

**Load data with read_csv**
```{r LoadData}
chem<-read_csv(here("Week6","Data","sgd_chem_data_HI.csv"))
```

### Create Table showing SGD parameters by Site

**Round decimals for mean and standard error**
```{r ChemTidy}
chem$Value_mean<-round(chem$Value_mean, digits = 3) # 3 decimal places
chem$Value_se<-round(chem$Value_se, digits = 2) # 2 decimal places
```

**Tidy data and create table**
```{r ChemTable}
chem_table<-chem %>% 
  unite(col = 'Mean +/- s.e.', Value_mean, Value_se, sep = " +/- ") %>% # combine mean and se columns
  select(-c(Value_variance, Value_sd)) %>%  # remove unnecessary columns
  pivot_wider(names_from = Site, values_from = 'Mean +/- s.e.') %>% # one column per site
  rename('Black Point  (mean +/- s.e.)' = BP, 'Wailupe  (mean +/- s.e.)' = W) %>%  # rename column headers
  mutate(Tide = ifelse(Tide == "High", "High Tide", "Low Tide")) %>% # relabel high and low tides
  select(-Variables) %>%  # remove variables column; will label variables below
  rename(Variables = Tide) %>% # rename tide column; variable names will be added later
  kbl(caption = "Table of measured SGD parameters at Black Point and Wailupe") %>% # create table with title
  kable_classic_2() %>% # format table style
  kable_styling(fixed_thead = T) %>% 
  row_spec(row = 0, bold = T, background = "#89689d", color = "#fff") %>%  # header specifications
  pack_rows(index = c("Nitrate+Nitrite"=2, # label variables above each set of High and Low Tide
                      "% SGD"=2, 
                      "pH"=2, 
                      "Phosphate"=2, 
                      "Salinity"=2, 
                      "Silicate"=2, 
                      "TA"=2, 
                      "Temperature in situ"=2),
            label_row_css = "background-color: #e69b99; color: #363434;") # set background color (consistent with PNW scale in plot) for row headers 
chem_table
```

### Create Figure showing SGD parameters by Site

**Tidy data and create plot**
```{r ChemPlot, fig.align = "center"}
pal<-pnw_palette("Starfish", 2) # set color palette for figure

chem_plot<-chem %>% 
  select(-(Value_variance:Value_se)) %>% # only retain mean values and standard error
  pivot_wider(names_from = Variables, values_from = Value_mean) %>% # pivot wider by variables
  rename('Nitrate+Nitrite' = NN, '% SGD' = percent_sgd, Temperature = Temp_in) %>% # rename col headings
  pivot_longer(cols = 'Nitrate+Nitrite':Temperature, names_to = "Variables", values_to = "MeanVal") %>% # retain temp as its own col
  mutate(Site = ifelse(Site == "BP", "Black Point", "Wailupe")) %>% 
  ggplot(aes(x = Tide,
             y = MeanVal,
             color = Site,
             fill = Site)) + 
  geom_col(position = "dodge", size = 0.4) +
  theme_bw() +
  labs(title = "Figure of mean SGD parameters at high and low tide",
       subtitle = "Locations: Black Point and Wailupe",
       y = "Mean Parameter Values",
       x = "Tide",
       caption = "Source: Lubarsky et al. 2018") +
  theme(axis.title = element_text(size = 14), axis.text = element_text(size = 10)) +
  scale_color_manual(values = pal) +
  scale_fill_manual(values = pal) +
  facet_wrap(~Variables, scales = "free")+
  ggsave(here("Week6","Output","ChemPlot.pdf"))
chem_plot  
```

