---
title: "Practicing with RMarkdown, Part 2"
author: "Danielle Barnas"
date: "3/3/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: journal
---

```{r}
knitr::opts_chunk$set(echo = TRUE, # the code will be displayed throughout doc
                      message = FALSE, # messages will not be displayed throughout doc
                      warning = FALSE, # warnings will not be displayed throughout doc
                      fig.path = "../Output/") # change figure path for all figures in doc
```

# Introduction
Today, we will use **Palmer Penguins** to manipulate figures in *RMarkdown*.  
We are using the "journal" theme.  
- Other themes can be found [here](https://www.datadreaming.org/post/r-markdown-theme-gallery/)

# Load Libraries

```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
library(memer) # today's totally awesome R package
```

**Figure options:**  
- out.width = "70%"
 - change plot size to 70% the normal size (or whatever relative size you want)  
- fig.width
 - figure width in "inches"  
- fig.height
 - figure height in "inches"  
- fig.align = "center"
 - location for figure in the document  
- fig.asp = 1.2
 - change the aspect ratio  
- fig.cap = "some caption"
 - figure caption where "the caption goes in quotes"  
- fig.path = " "
 - because our current file is saved in Scripts, fig.path would automatically save in Scripts, but to save in a different folder, use "../Folder/" where the ../ goes up one folder in the directory.  

**We can also label a code chunk by putting a name after the r and before the comma**  
- Good practice to give all our code chunks a different name  
- Can never have a repeated name, else we will get an error  
- If you are saving the figure, the code chunk name will be the name of the figure or figures.  

# Make a plot

```{r PengFig, out.width="80%", fig.align="center", fig.cap="This is a figure of penguin flippers", fig.path = "../Output/", echo = FALSE}
PenguinFigure<-penguins %>% 
  ggplot(aes(x = species, y = flipper_length_mm, colour = island)) + 
  geom_boxplot()

PenguinFigure
```

# Make a table

| Time | Session | Topic |
|:-----|:-------:|------:|
| _left_ | _center_ | _right_ |
| 01:00  | 1       | Anatomy |
| 01:50  |         | *Break* |
| 02:00  | 2       | Tables  |
| 02:45  |         | *Break* |

Line one contains table headers  
Line two separates table headers from table contents  
- Colon indicates alignment: left = left, left and right = center, right = right  

```{r}
# create a crude table
penguins %>% 
  group_by(species) %>% 
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>% 
  kbl() %>%  # tidy up the table
  kable_classic() %>%  # adds nice thick lines, slightly more journal quality
  row_spec(2, # can do some specific command to a specific row (in this case, row 2)
           bold = TRUE, # bold text
           color = "white", # text color is white
           background = "red") # highlight color is red
```

Today's Totally Awesome R Package  
- [memer](https://github.com/sctyner/memer/)

```{r}
library(memer)
# meme's automatically save
meme_get("DistractedBf") %>% 
  meme_text_distbf("Silly R package", "BIOL 551 students", "actual coding")

meme_get("DosEquisMan") %>% 
  meme_text_top("I don't always like coding", size = 28) %>% 
  meme_text_bottom("But when I do\nIt's always today's totally awesome R package", size = 18)
```



