---
title: "Getting Help with R"
author: "Danielle Barnas"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```


## Googling
- Use the package and function if known  
- Put .R or R software to make more specific  
- use 'dataframe' rather than 'spreadsheet'  
- CRAN Task Views  
  - can help you find a package through keywords if you don't know the package


## Troubleshooting Script
- Sleep!  
- Look for red ex's on the side of code or red squiggly lines withitn your code, indicating syntax errors  
- Many messages R will print to let you know the problem  
  - Error: There's a fatal error in your code that prevented it from being run through successfully. You need to fix it for the code to run  
  - Warning: non-fatal errors (don't stop code from running, but this is a potential problem)  
  - Message: Just a note about your code  


## Supportive R Communities
- Stack Overflow  
- Twitter  
  - #Rstats

## Helpful online guides
- https://ivelasq.rbind.io/blog/other-geoms/#waffle-charts  
- r-graph-gallery.com  
- https://ivelasq.rbind.io/blog/other-geoms/#waffle-charts

## Make a reproducible example to ask for help
- Re: StackOverflow
```{r}
# install.packages('reprex') # reproducible example
# install.packages('datapasta') # copy and paste
# install.packages('styler') # copy and paste in style

library(reprex)
library(datapasta)
library(styler)
```


### Reprex
```{r, eval = FALSE}
library(tidyverse)
mpg %>%
  ggplot(aes(x = displ, y = hwy))%>%
  geom_point(aes(color = class))
```
1. Highlight code  
1. Addins > reprex selection  
1. Addins > render reprex...
  1. Select Current Selection
  1. Select a Target venue
  1. Select Append session info and Preview HTML


### Datapasta
```{r, eval = FALSE}
lat	long	star_no
33.548	-117.805	10
35.534	-121.083	1
39.503	-123.743	25
32.863	-117.24	22
33.46	-117.671	8
```

```{r}
tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L
  )

```

1. copy and paste excel table data
1. addins > Datapasta > Paste as tribble (or other format as relevant)



### Do's and Don't of Reprex
1. Use the smallest, simplest, most built-in data possible
  1. e.g. using datapasta
1. Include commands on a strict "need to run" basis
  1. ruthlessly strip out anything unrelated to the specific matter at hand
  1. Include every single command that is required, e.g. loading specific packages as needed
1. Pack it in, pack it out, and don't take liberties with other people's computers. You are asking people to run this code!
  1. Don't start with rm(list = ls()). It is anti-social to clobber other people's workspaces
  1. Don't start with setwd() because it won't work on anyone else's computer
  1. Don't mask built-in functions, i.e. don't define a new function named c or mean
  1. If you create files, delete them when you're done
1. Use good coding style, e.g. include comments
1. Make sure you run your code first, don't fake it
1. Should be easy for others to digest, so they don't necessariliy have to run it
1. Should be easy for others to copy + paste + run, if and only if they so choose.


## Totally awesome R package
- ubeR
  - call a ride through R
  
  
## Lab:
Create a simple code with an error(s)

```{r, eval = FALSE}
library(tidyverse)

head(mtcars)

ggplot(data = mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = cyl) +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Horsepower", y = "MPG", color = "# Cylinders", 
       title = "Scatter Plot of HP vs MPG") +
  scale_color_viridis_d()
```






