---
title: "Week 7 Map Practice"
author: "Danielle Barnas"
date: "3/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, # ALWAYS show the code
                      warning = FALSE, # NEVER show the warnings
                      message = FALSE) # NEVER show messages
```

# Load Libraries
```{r}
# Load Libraries
library(tidyverse)
library(here)
library(maps)
library(mapdata)
library(mapproj)
```


# Bring in Data
```{r}
## Bring in Data
# data on population in California by county
popdata<-read_csv(here("Week7","Data","CApopdata.csv"))
# data on number of seastars at different field sites
stars<-read_csv(here("Week7","Data","stars.csv"))
```


# {maps} Package
map_data("location") is the function used to pull out whatever base layer that you want  
```{r}
## {maps} package
# get data fro the entire world
world<-map_data("world")
head(world)

usa<-map_data("usa") # creates a polygon at the country level
head(usa)

italy<-map_data("italy")
head(italy)

states<-map_data("state") # create polygons at state level
head(states)

counties<-map_data("county") # creates polygons at the county level
head(counties)

```


# Structure of data
- group: very important for ggplot to know how to structure the map  
  - in the map_data data frame there is a group number for the different polygons that are meant to stay together  
  - otherwise there's just a bunch of weird lines


# Make a map of the world
- can use different projections to view the map  
  - https://rdrr.io/cran/mapproj/man/mapproject.html  
```{r}
# absolute basic
ggplot()+
  geom_polygon(data = world, aes(x = long, y = lat, group = group))

# with additions
ggplot() +
  geom_polygon(data = world, 
               aes(x = long, 
                   y = lat, 
                   group = group,
                   fill = region), # fill in colours of map by region
               color = "black") + # colour outside of aes because not manipulating the data, just making outlines black
  theme_minimal() + # gets rid of box around the figure
  guides(fill = F) +
  theme(panel.background = element_rect(fill = "lightblue")) + # changes background panel to be light blue. element_rect is the rectangular element
  coord_map(projection = "mercator", # also showed "sinusoidal"
            xlim = c(-180,180)) # specific about using the whole world, but not outside those limits
```


# Make a map of California
```{r}
CA_data<-states %>% 
  filter(region == "california")
head(CA_data)
```

## Think, Pair, Share
- Make a simple map of just california
```{r}
CA_data<-states %>% 
  filter(region == "california")

ggplot() +
  geom_polygon(data = CA_data, 
               aes(x = long, 
                   y = lat, 
                   group = group), # fill in colours of map by region
               color = "black") + # colour outside of aes because not manipulating the data, just making outlines black
  theme_void() + # just the map and nothing else (removes lat and long)
  guides(fill = F) +
  theme(panel.background = element_rect(fill = "lightblue")) + 
  coord_map()  # "albers" is good for California
```

# Adding multiple layers of data
```{r}
CApop_county<- popdata %>% 
  select("subregion" = County, Population) %>% # rename and select relevant columns
  inner_join(counties) %>% # join popdata to counties
  filter(region == "california") # only want CA data
```

# Map of CA population by county
```{r}
ggplot() +
  geom_polygon(data = CApop_county,
               aes(x = long,
                   y = lat,
                   group = group,
                   fill = Population),
               color = "black") + 
  geom_point(data = stars, #add a point at seastar sampling locations
             aes(x = long,
                 y = lat,
                 size = star_no)) + # the size of the point relates to the numebr of seastars found at each location
  coord_map() +
  #coord_map(projection = "albers",parameters = c(-38, 118)) + 
  theme_void() + 
  scale_fill_gradient(trans = "log10") + # change to the fill gradient. transformed to log10. Colors are more easily distinguishable 
  labs(size = "# stars/m2") + # clean up legend
  ggsave(here("Week7","Output","CApop.pdf"))
```


Today's Totally Awesome R Package
```{r}
library(spongebob)

tospongebob("No this is Patrick")

spongebobsay(what = "No this is Patrick")
```


