---
title: "Iterative Coding"
author: "Danielle Barnas"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Libraries
```{r}
library(tidyverse)
library(here)

```


# For Loops

for(index in sequence) {  
command to repeat  
}  

## Simple for loop

- Start with the basic code, only having it run once.  
- Once it works, then add in the iterations

```{r}
print(paste("The year is", 2000))
```

Put it in a for loop
```{r}
years <- (2015:2021)

for(i in years){ # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}
```

At this point, there's nothing new in our environment, we've just printed over and over.  Usually we want to have some kind of output.  
- First we need to tell R how much space we need to store data  
- Create an empty dataframe to pre-allocate space  

```{r}
# Pre-allocate space for the for loop
# empty matrix
year_data <- data.frame(matrix(ncol = 2, nrow = length(years))) # number of rows = length of the vector 'years'

colnames(year_data) <- c("year", "year_name")

year_data
```

For loops tend to be slower if you don't tell R beforehand how much space to allocate.  If R has to keep building and building through the for loop, then it can be slower.

```{r}
for(i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i] <- paste("The year is", years[i])
}

year_data
```


```{r}
for(i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i] <- paste("The year is", years[i]) # put phrase in row i
  year_data$year[i] <- years[i] # loop over year
}

year_data
```

# Using loops to read in multiple .csv files


Read in one file
```{r}
testdata <- read_csv(here("Week13", "Data", "cond_data", "011521_CT316_1pcal.csv"))
```

List files in a directory
```{r}
# point to the location on the computer of the folder
CondPath <- here("Week13","Data","cond_data")

# list all the files in that path with a specific pattern
# In ths case we are looking for everything that has a .csv in the filename

# you can use regrex to be more specific if you are looking for certain patterns
files <- dir(path = CondPath, pattern = ".csv")
```

Pre-allocate space for the loop

Let's calculate the mean temp and salinity from each file and save it

```{r}
cond_data <- data.frame(matrix(nrow = length(files), ncol = 3))

colnames(cond_data) <- c("filename", "mean_temp", "mean_salinity")
cond_data
```

for loop
```{r}
raw_data <- read_csv(paste0(CondPath,"/",files[1])) # test by reading in the first file

head(raw_data)
```
Test code for mean temperature
```{r}
mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean
mean_temp
```

# Turn it into a for loop
```{r}
for (i in 1:length(files)) { # loop over the number of files (1:3)
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  
  cond_data$filename[i] <- files[i] # add file name to the filename column
}

cond_data
```

Add in means
```{r}
for (i in 1:length(files)) { # loop over the number of files (1:3)
  raw_data <- read_csv(paste0(CondPath,"/",files[i]))
  
  cond_data$filename[i] <- files[i] # add file name to the filename column
  cond_data$mean_temp[i] <- mean(raw_data$Temperature, na.rm = TRUE)
  cond_data$mean_salinity[i] <- mean(raw_data$Salinity, na.rm = TRUE)
}

cond_data
```

# Map functions

the pattern of looping over a vector, doing somethign to each element and saving the results is so common that the purrr package provides a family of functions to do it for you

map() makes a list  
map_lgl() makes a logical vector
map_int() makes an integer vector
map_dbl() makes a double vector
map_chr() makes a character vector
map_df() makes a dataframe

## Simple example

1 task in 3 ways

Calculate the mean from a random set of numbers and do it 10 times  
Create a vector from 1:10
```{r}
1:10 # a vector from 1 to 10 (since we are doing this 10 times)
```

for each time 1:10 make a vector of 15 random numbers based on a normal distribution

```{r}
1:10 %>% # a vector from 1 to 10 (since we are doing this 10 times
  map(rnorm, n = 15) %>% # calculate 15 random numbers based on a normal distribution
  map_dbl(mean) # calculate the mean. It is now a vector which is type "double"
```


Same thing, different notation..  
Make your own function
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15,x)) %>%  # make your own function
  map_dbl(mean)
```

Use a formula when you want to change the arguments within the function

```{r}
1:10 %>% 
  map(~ rnorm(15, .x)) %>%  # changes the argumnents inside the function
  map_dbl(mean)
```

# Bring in files using purrr instead of a for loop

find the files
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_13", "data", "cond_data")
files <- dir(path = CondPath,pattern = ".csv")
```


Or, get the full file name in one less step by doing this

```{r}
files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```


Read in the files, and bind them together
```{r}
data <- files %>% 
  set_names() %>% 
  map_df(read_csv, .id = "filename")
```

Calculate means
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") %>% # map everything to a dataframe and put the id in a column called filename
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))

data
```

# Totally awesome R package

{gganatogram}

Use R for anatomy

```{r, eval = FALSE}
# Install
Sys.setenv(R_REMOTES_NO_ERRORS_FROM_WARNINGS="true")
## You might get a weird warning from this package about converting a warning to error.  This will get rid of that
remotes::install_github("jespermaag/gganatogram")
```




# Lab

Bring in Jenn's files using both a for loop and then also purrr map functions in your script.

Don't need to save any output, but need to make sure our code can run:

The goal is to bring in all 4 files and calculate the mean and standard deviation of both temperature (Temp.C) and light (Intensity.lux) for each tide pool.


## Standard for loop
```{r}
file.path <- here("Week13","Data","homework") # file path
files <- dir(path = file.path,pattern = ".csv") # find files in path; one file per tidepool
files

# create empty matrix for storing data
pool.data <- data.frame(matrix(nrow = length(files), ncol = 5))
colnames(pool.data) <- c("filename", "mean_temp", "sd_temp","mean_lux", "sd_lux")

for(i in 1:length(files)){ # 1:4
  mydata <- read_csv(paste0(file.path,"/",files[i]))
  pool.data$filename[i] <- files[i]
  pool.data$mean_temp[i] <- mean(mydata$Temp.C, na.rm = TRUE)
  pool.data$sd_temp[i] <- sd(mydata$Temp.C, na.rm = TRUE)
  pool.data$mean_lux[i] <- mean(mydata$Intensity.lux, na.rm = TRUE)
  pool.data$sd_lux[i] <- sd(mydata$Intensity.lux, na.rm = TRUE)
}
pool.data
```

## purrr

```{r}
file.path <- here("Week13","Data","homework") # file path
files <- dir(path = file.path,pattern = ".csv", full.names = T) # find files in path; one file per tidepool

data<-files %>%
  set_names() %>% # sets the id of each list to the file name
  map_df(read_csv, .id = "filename") %>% # map everything to a dataframe and put the id (from set_names()) in a column called filename
  group_by(filename) %>% # effectively groups by tide pool
  summarise(mean_temp = mean(Temp.C, na.rm = TRUE),
            sd_temp = mean(Temp.C,na.rm = TRUE),
            mean_lux = mean(Intensity.lux, na.rm = TRUE),
            sd_lux = mean(Intensity.lux, na.rm = TRUE))
data
```



