---
title: "Modeling"
author: "Danielle Barnas, Group 1"
date: "4/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = F,
                      message = F)
```

# Libraries
```{r}
library(tidyverse)
library(here)
library(palmerpenguins)
library(broom)
library(performance)
library(modelsummary)
library(tidymodels)
```



# Anatomy of a basic linear model

## Simple Linear Model
mod<-lm(y ~ x, data = df)

lm = linear model, y = dependent variable, x = independent variable

"y is a function of x"

## Multiple regression
mod<-lm(y~x1 + x2, data = df)

## Interaction term
mod<-lm(y~x1*x2, data = df) the * will compute x1+x2+x1:x2


## Working with Palmer Penguins
```{r}
Peng_mod<-lm(bill_length_mm ~ bill_depth_mm*species, data = penguins)
```



## Check model assumptions with performance package
```{r}
check_model(Peng_mod)
```

## View results: base R
ANOVA Table
```{r}
anova(Peng_mod)
```

Coefficients (effect size) with error
```{r}
summary(Peng_mod)
```

## Broom package cleans up model output data
- tidy  
- glance  
- augment  

```{r}
# Tidy coefficients
coeffs <- tidy(Peng_mod)
coeffs
```

```{r}
# Tidy r2, etc; puts all in one row
results <- glance(Peng_mod)
results
```

```{r}
# Tidy residuals, etc; same number of rows as original dataframe
resid_fitted <- augment(Peng_mod)
resid_fitted
```

## Results in modelsummary

Creates tables and plots to summarize statistical models and data in R


Compare Peng_mod with one that does not have hte species as an interaction term
```{r}
# New model
Peng_mod_noX<-lm(bill_length_mm ~ bill_depth_mm, data = penguins)
#Make a list of models and name them
models<-list("Model with interaction" = Peng_mod,
             "Model with no interaction" = Peng_mod_noX)
#Save the results as a .docx
modelsummary(models, output = here("Week13","Output","table.docx"))
```

## Modelplot

Canned coefficient modelplots
```{r}
modelplot(models) +
  labs(x = 'Coefficients',
       y = 'Term names')
```

## Many models with purrr, dplyr, and broom

Let's say you want to plot and compare lots of different models at the same time and view the results

First, createa a list for the data by each species.  We do this using nest().  We are going to nest the data by species

Nest takes all the data and puts it into a set of lists.  Here we're saying "nest all the data except for species"

```{r}
models <- penguins %>% 
  ungroup() %>% # the penguin data are grouped so we need to ungroup them
  nest(-species) # nest all teh data by species

m
models$data
```


## map a model to each of the groups in the list

```{r}
models<- penguins %>%
  ungroup()%>% # the penguin data are grouped so we need to ungroup them
  nest(-species) %>% # nest all the data by species 
  mutate(fit = map(data, ~lm(bill_length_mm~body_mass_g, data = .)))

models

models$fit # shows you each of the 3 models
```


View the results. First, let's mutate the models list so that we have a tidy coefficient dataframe (using tidy()) and a tidy model results dataframe (using glance())
```{r}
results<-models %>%
   mutate(coeffs = map(fit, tidy), # look at the coefficients
          modelresults = map(fit, glance))  # R2 and others
results
```

Next, select what we want to show and unnest it to bring it back to a dataframe

```{r}
results<-models %>%
   mutate(coeffs = map(fit, tidy), # look at the coefficients
          modelresults = map(fit, glance)) %>% # R2 and others 
   select(species, coeffs, modelresults) %>% # only keep the results
   unnest() # put it back in a dataframe and specify which columns to unnest

View(results)
```


## Other very common stats packages

- stats: General (lm)and generalized (glm) linear models (already loaded with base R)  
- lmer : mixed effects models  
- lmerTest' : getting results from lmer  
- nlme : non-linear mixed effects models  
- mgcv, gam : generalized additive models  
- brms, rstan, and many more : Bayesian modeling  
- lavaan, peicewiseSEM : Structural Equation Models  
- rpart, randomForest, xgboost, and more : Machine learning models  


# Tidymodels
Uses tidyverse lenguage to run models

Start by specifying the *functional form* using the parsnip package.  In our case, we will use a linear regression which is coded like:

linear_reg()

Then set the *engine* for what type of linear regression we are modeling.  For example, we could use an OLS regression or Bayesian or several other options.

```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm")

lm_mod
```

Next we add the model fit
```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins)

lm_mod
```


Lastly, we add the tidy it. And now we can pipe this into plots, etc. Nice, tidy way to model.
```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins) %>%
  tidy()

lm_mod
```


## Pipe to a plot
```{r}
lm_mod<-linear_reg() %>%
  set_engine("lm") %>%
  fit(bill_length_mm ~ bill_depth_mm*species, data = penguins) %>%
  tidy() %>%
  ggplot()+
    geom_point(aes(x = term, y = estimate))+
    geom_errorbar(aes(x = term, ymin = estimate-std.error,
                      ymax = estimate+std.error), width = 0.1 )+
  coord_flip()

lm_mod
```

# Today's Totally Awesome R Package

## Pushoverr
```{r}
library(pushoverr)
```


# Final Presentation 

- Go back to recording in Lecture 4, second video, 20 minutes in
- Read over rubric
- If you do markdown, make it beautiful.
- The goal is to make everything publication quality
  - figures
  - markdown
  - website
  - shiny app
  - xarigan
  - etc. make it fancy
- Make the new repository in the Biol551 Organization
- Look at specific expectations for the README.md file
- Everything is due by 9AM on May 10







