---
title: "Creating functions"
author: "Danielle Barnas"
date: "3/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

Why functions?

Create a table
```{r}
library(tidyverse)

df<-tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

head(df)
```

Mutate - try to avoid mistakes
```{r}
df<-df %>% 
  mutate(a = (a-min(a, na.rm = T))/(max(a, na.rm = T)-min(a, na.rm = T)))

# Copy and paste for a through d

df<-df %>% 
  mutate(a = (a-min(a, na.rm = T))/(max(a, na.rm = T)-min(a, na.rm = T))) %>%
  mutate(b = (b-min(b, na.rm = T))/(max(b, na.rm = T)-min(b, na.rm = T))) %>%
  mutate(c = (c-min(c, na.rm = T))/(max(c, na.rm = T)-min(c, na.rm = T))) %>%
  mutate(d = (d-min(d, na.rm = T))/(max(d, na.rm = T)-min(d, na.rm = T)))
```


We can write a function for this
1. Pick a useful name for the function  
2. List the inputs, or arguments, to the function inside the function  
3. Place the code in the body of the function  
4. Return tells us what values we want returned when you run the function  
```{r}
rescale01<-function(x) {
  value<-(x-min(x, na.rm = T))/(max(x, na.rm = T)-min(x, na.rm = T))
  return(value)
}
```


Test the function
```{r}
df %>% 
  mutate(a = rescale01(a),
         b = rescale01(b),
         c = rescale01(c),
         d = rescale01(d))
```
- Useful to put the functions at the top of your script so that the function is available throughout the script.  


Make a function to convert degrees F to degrees C
```{r}
# temp_C <- (temp_F - 32) * 5 / 9

f_to_c <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}
```

Test function
```{r}
f_to_c(32)
f_to_c(212)
```


Create a function to convert celcius to Kelvin:  
Kelvin = calcius + 273.15
```{r}
c_to_k <- function(temp_c) {
  temp_k <- temp_c + 273.15
  return(temp_k)
}
```

Test function
```{r}
c_to_k(20)
```


Making plots into a function
```{r}
library(palmerpenguins)
library(PNWColors)

pal<-pnw_palette("Lake",3,type = "discrete")

ggplot(penguins, 
       aes(x = body_mass_g,
           y = bill_length_mm,
           color = island)) + 
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()


# Now use the plot to make a function
myplot<-function(data,x,y) {
  pal<-pnw_palette("Lake",3,type = "discrete")

ggplot(data, 
       aes(x = x,
           y = y,
           color = island)) + 
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

```

Test function
```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)

# Doesn't work!  Beacse we don't already have body mass or bill length in our environment
```

Use a "curly-curly" {{}} to help assign variables that are column names in dataframes.

Try again with the curly-curlies
```{r}
myplot<-function(data, x, y){ 
pal<-pnw_palette("Lake", 3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
```

Test again
```{r}
myplot(data = penguins, x = body_mass_g, y = bill_length_mm)
myplot(data = penguins, x = body_mass_g, y = flipper_length_mm)
```

Adding defaults
```{r}
myplot<-function(data = penguins, x, y){
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

myplot(x = body_mass_g, y = flipper_length_mm)
```


Layering the plot (like a normal ggplot)
```{r}
myplot(x = body_mass_g, y = flipper_length_mm)+
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)")
```

Add an if-else statement for more flexibility
```{r}
a <- 4
b <- 5

# Suppose that if a > b then f should be = 20, else f should be equal to 10
if(a > b) { # my question
  f <- 20 # if it is true give me answer 1
} else { # else give me answer 2
  f <- 10
}

f # gives me f's value, which is 10
```


Back to plotting (using if-else)
```{r}
# general plot function
myplot<-function(data = penguins, x, y ,lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}

# Add an if-else statement
myplot<-function(data = penguins, x, y, lines=TRUE ){ # add new argument for lines
pal<-pnw_palette("Lake",3, type = "discrete") # my color palette 
if(lines==TRUE){
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  geom_smooth(method = "lm")+ # add a linear model
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
else{
ggplot(data, aes(x = {{x}}, y = {{y}} , color = island))+
  geom_point()+
  scale_color_manual("Island", values=pal)+   # use pretty colors and change the legend title
  theme_bw()
}
}
```

Test it
```{r}
# with plot lines
myplot(x = body_mass_g, y = flipper_length_mm)

# without plot lines
myplot(x = body_mass_g, y = flipper_length_mm, lines = FALSE)
```


Totally awesome R package of the day
install.packages("fortunes")
```{r}
library(fortunes)
fortune()
```






