---
title: "Lab - Creating functions"
author: "Danielle Barnas"
date: "3/24/2021"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


# Assignment:  
Use any of the datasets that we have used in class and make two useful functions. One that is a plot and one that is not. Turn in your functions using RMarkdown with at least 2 examples each showing its flexibility. Due, as always, at 9am on Monday.


# Load Libraries
```{r}
library(tidyverse)
library(palmerpenguins)
library(PNWColors)
library(ggalt)
library(here)
```


# Glimpse Data
```{r}
glimpse(penguins)
```


# Create function (no plot)
```{r}
# Create function for some biomass:length ratio
BM_ratio <- function(data, body_mass, length) {
  ratio <- data %>% 
    mutate(body_mass_ratio = {{body_mass}} / {{length}}) %>%  # create new ratio column
    relocate(body_mass_ratio, .before = {{body_mass}}) # move from end of dataframe so it doesn't get cut off in the output
  return(ratio)
}
```

### Test function
```{r}
# body mass / bill length ratio
BM_ratio(data = penguins, body_mass = body_mass_g, length = bill_length_mm)

# body mass / flipper length ratio
BM_ratio(data = penguins, body_mass = body_mass_g, length = flipper_length_mm)
```


# Create function (plot)
```{r}
myplot<-function(data, x, y, group){ # set variables
  group_distinct<- data %>% #isolate distinct variables in the group
    distinct({{group}})
  pal<-pnw_palette("Sunset2", nrow(group_distinct)) # color palette for plot, nrow gives number of distinct variables in the group
  data<-data %>% 
    drop_na({{group}}) # remove na values from the chosen group
  ggplot(data, # assign dataframe
         aes(x = {{x}}, y = {{y}}, # assign variables
             color = {{group}}, # color points by group
             fill = {{group}}, # fill circles behind points
             shape = {{group}})) + # each group variable has a different shape
    geom_point() + # scatter plot
    geom_encircle(alpha = 0.2, show.legend = FALSE) + # circle data with shaded background
    scale_color_manual(values = pal) + # set palette
    scale_fill_manual(values = pal) + # set palette
    theme_bw() +
    theme(axis.title = element_text(size = 12),
          axis.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          legend.text = element_text(size = 10))
}
```

### Test function
```{r}
myplot(data = penguins, 
       x = body_mass_g, 
       y = bill_length_mm,
       group = species)
```

### Test function with add-ons:

### Example 1
```{r}
myplot(data = penguins, 
       x = body_mass_g, 
       y = bill_length_mm,
       group = species) +
  labs(x = "Body mass (g)",
       y = "Bill length (mm)",
       title = "Palmer penguin body mass and bill length distribution by species") +
  ggsave(here("Week8","Output","Lab_function_plot_Ex1.pdf"), width = 7, height = 7)
```

### Example 2
```{r}
myplot(data = penguins, 
       x = body_mass_g, 
       y = flipper_length_mm,
       group = island) +
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)",
       title = "Palmer penguin body mass and bill length distribution by island") +
  ggsave(here("Week8","Output","Lab_function_plot_Ex2.pdf"), width = 7, height = 7)
```











